### 2013-06-17 gapoly.Rmd
1st commit

GA Poly
========================================================

```{r, fig.width=12}
library(genalg)

x1 <- seq(0, 4 * pi, length.out = 201)
x2 <- seq(0, 4 * pi, length.out = 201)

# a latent variable, with a trick expression and some noise
y <- 2 * sin(x1) + cos(2 * x2) + 0.1 * x1 * x2 + 0.1 * rnorm(201)

# Caution: there are **ONLY** 201 observations (and not 201*201)
observations <- data.frame(y = y, x1 = x1, x2 = x2)

##############

make.component <- function(bits, n.vars) {
  vars <- paste0("x",1:n.vars)
  elements <- split(bits, ceiling(seq_along(bits)/(length(bits)/n.vars)))
  str <- ""
  for(i in 1:length(elements)) {
    power <- length(elements[[i]][elements[[i]]==1])
    if (power == 1)
      if (str=="")
        str = vars[i]
      else  
        str = paste0(str, "*", vars[i])
    if (power > 1)
      if (str=="")
        str = paste0(vars[i], "^", power)
      else  
        str = paste0(str, "*", vars[i], "^", power)
  }
  str
}

make.formula <- function(bits, n.vars, n.compts) {
  components <- split(bits, ceiling(seq_along(bits)/(length(bits)/n.compts)))
  str <- ""
  for(i in 1:length(components)) {
    comp <- make.component(components[[i]], n.vars)
    if (comp!="")
      if (str=="")
        str = paste0("I(",comp,")")
    else
        str = paste0(str," + I(", comp ,")")
  }
  str
}

############################

max.degree <- 3
n.vars     <- 3  # x1, x2,... 
n.compts   <- 4  # poly = compt1 + compt2 + ...

set.seed(101)
bits <- sample(c(0,1), max.degree*n.vars*n.compts, replace=TRUE)
paste0(bits, collapse = "") # just to show as string

##############################

x1 <- seq(0, 8 * pi, length.out = 201)/pi
x2 <- sin(x1)^2
x3 <- abs(cos(x2))
y  <- x1*x2 - x1*x3 + x3^3 + rnorm(201,0,.15) # a latent variable with some noise
df <- data.frame(y=y,x1=x1,x2=x2,x3=x3) # make data frame

evalFuncFactory <- function(df, n.vars, n.compts) {

  function(bits) {
    formula <- paste0("y ~ ", make.formula(bits, n.vars, n.compts))
    model <- lm(formula, data=df)
    return( sqrt(mean(residuals(model)^2)) )
  }
}

#############################

GAmodel <- rbga.bin(size = max.degree*n.vars*n.compts, 
                    popSize = 100, 
                    iters = 100, 
                    mutationChance = 0.01, 
                    elitism = TRUE, 
                    evalFunc = evalFuncFactory(df, n.vars, n.compts))

cat(summary.rbga(GAmodel))

best.solution <- GAmodel$population[1,]
best.formula <- paste0("y ~ ", make.formula(best.solution, n.vars, n.compts))
model <- lm(best.formula, data=df)
coef(model)
plot(df$x1,df$y,type="l")
title(best.formula)
y1<-predict(model,df[,-1])
lines(x1,y1,col="red") # approx
```