### 2013-06-17 gapoly.Rmd
changed dataframe eg

GA Poly
========================================================

GA package use based on [http://www.r-bloggers.com/genetic-algorithms-a-simple-r-example/](http://www.r-bloggers.com/genetic-algorithms-a-simple-r-example/)

<hr>

Herein we code a polynomial based on variables $x_1, x_2 \ldots x_n$ into a binary vector (ie, made of 0s and 1s).

To achieve this we need to know three values in advance:
+ number of different variables
+ maximum number of different components (being a polynomial a sum of components)
+ maximum degree for any given variable

```{r}
n.vars     <- 3  # x1, x2,... 
max.compts <- 4  # poly = compt1 + compt2 + ...
max.degree <- 3
```

The chromosome coding will as follow:
+ the chromosome is split into `max.compts` sets of bits of equal size (a component)
+ each component is split into `n.vars` sets of `max.degree` size each (a variable)
+ for each variable, the number of 1s gives the variable's degree (this means that a given polynomial may have more than one possible representation)

Eg: consider polynomial $x_1^2 \times x_3 + x_3^3 + x_1 \times x_2$

One possible coding -- for the values in the previous code snippet -- would be:

$$011,000,001;000,000,111;001,001,000;000,000,000$$

(the semicolons separate components, the commas seperate variables)

The next functions transform the binary vector into a polynomial formula that `lm()` understands:

```{r}
make.component <- function(bits, n.vars) {
  vars <- paste0("x",1:n.vars)
  elements <- split(bits, ceiling(seq_along(bits)/(length(bits)/n.vars)))
  str <- ""
  for(i in 1:length(elements)) {
    power <- length(elements[[i]][elements[[i]]==1])
    if (power == 1)
      if (str=="")
        str = vars[i]
      else  
        str = paste0(str, "*", vars[i])
    if (power > 1)
      if (str=="")
        str = paste0(vars[i], "^", power)
      else  
        str = paste0(str, "*", vars[i], "^", power)
  }
  str
}

make.formula <- function(bits, n.vars, max.compts) {
  components <- split(bits, ceiling(seq_along(bits)/(length(bits)/max.compts)))
  str <- ""
  for(i in 1:length(components)) {
    comp <- make.component(components[[i]], n.vars)
    if (comp!="")
      if (str=="")
        str = paste0("I(",comp,")")
    else
        str = paste0(str," + I(", comp ,")")
  }
  str
}
```

Let's check the previous eg, $x_1^2 \times x_3 + x_3^3 + x_1 \times x_2$:

```{r}
code.eg <- c(0,1,1,0,0,0,0,0,1,0,0,0,0,0,0,1,1,1,0,0,1,0,0,1,0,0,0,0,0,0,0,0,0,0,0,0)
paste0(code.eg, collapse = "") # just to show as string
make.formula(code.eg, n.vars, max.compts)
```

Let's create a dataframe for testing. Notice that the data frame columns name must be called $y$ for the output value, and $x_i$ for the input values:

```{r, fig.width=12}
set.seed(102)
x1 <- seq(0, 8 * pi, length.out = 201)/pi
x2 <- sin(x1)^2
x3 <- abs(cos(x1))
y  <- x1*x2 - x1*x3 + x3^3 + rnorm(201,0,.15) # a latent variable with some noise
df <- data.frame(y=y,x1=x1,x2=x2,x3=x3) # make data frame
```

This binary coding is useful to use the binary GA `rbga.bin()` from package `genalg` (see below):

```{r}
library(genalg)

# This is a factory for the evaluation function that the GA function
# needs to execute.

# The factory is a closure that includes the dataframe, the number of
# variable and the max number of componentes, all required information
# to perform the polynomial regression and then compute the residuals' rsme 
# which will be the fitness of the respective chromosome
evalFuncFactory <- function(df, n.vars, max.compts) {

  function(chromosome) {
    formula <- paste0("y ~ ", make.formula(chromosome, n.vars, max.compts))
    model <- lm(formula, data=df)
    # rbga.bin() minimizes, so the rmse will be smaller as the interations advance
    return( sqrt(mean(residuals(model)^2)) ) 
  }
}

#################################
# Unleash the chromosomes of war!
#################################
GAmodel <- rbga.bin(size = max.degree*n.vars*max.compts, 
                    popSize = 150, 
                    iters = 100, 
                    mutationChance = 0.01, 
                    elitism = TRUE, 
                    evalFunc = evalFuncFactory(df, n.vars, max.compts))

cat(summary.rbga(GAmodel))
best.solution <- GAmodel$population[1,]
best.formula <- paste0("y ~ ", make.formula(best.solution, n.vars, max.compts))
model <- lm(best.formula, data=df)
coef(model)
plot(df$x1,df$y,type="l")
title(best.formula)
y1<-predict(model,df[,-1])
lines(x1,y1,col="red") # approx
```