### 2013-06-28 PaperDraft.Rmd
The first steps towards a paper.

Learning Polynomial Regression with Genetic Algorithms
======================================================

Francisco Coelho |  | João Neto
:---------------:|-----|:---------:
[`fc@di.uevora.pt`](mailto://fc@di.uevora.pt)| |[`jpn@di.fc.ul.pt`](mailto://jpn@di.fc.ul.pt)

Introduction
------------

_intro text_

### State of the Art

Em [1] vimos que...

Section 1
---------

References
----------

1. Hofwing, Magnus, Niclas Strömberg, and Martin Tapankov. "Optimal Polynomial Regression Models by using a Genetic Algorithm." Proceedings of the Second International Conference on Soft ComputingTechnology in Civil, Structural and Environmental Engineering Conference,(Crete, Greece), 2011009. 2011.

1. Wu, Chih-Hung, Gwo-Hshiung Tzeng, and Rong-Ho Lin. "A Novel hybrid genetic algorithm for kernel function and parameter optimization in support vector regression." Expert Systems with Applications 36.3 (2009): 4725-4735.

1. Gálvez, Akemi, Andrés Iglesias, and Jaime Puig-Pey. "Iterative two-step genetic-algorithm-based method for efficient polynomial B-spline surface reconstruction." Information Sciences 182.1 (2012): 56-76.

1. Ghai, Dhruva, Saraju P. Mohanty, and Garima Thakral. "Fast Analog Design Optimization using Regression based Modeling and Genetic Algorithm: A Nano-CMOS VCO Case Study." Proceedings of the 14th IEEE International Symposium on Quality Electronic Design (ISQED). 2013.

1. Rezania, Mohammad, Akbar A. Javadi, and Orazio Giustolisi. "Evaluation of liquefaction potential based on CPT results using evolutionary polynomial regression." Computers and Geotechnics 37.1 (2010): 82-92.

1. Zain, Azlan Mohd, Habibollah Haron, and Safian Sharif. "Genetic algorithm and simulated annealing to estimate optimal process parameters of the abrasive waterjet machining." Engineering with computers 27.3 (2011): 251-259.

1. Garg, A., and K. Tai. "Comparison of regression analysis, artificial neural network and genetic programming in handling the multicollinearity problem." Modelling, Identification & Control (ICMIC), 2012 Proceedings of International Conference on. IEEE, 2012.