### 2013-07-02 tests_test.Rmd

```{r, message = FALSE, tidy = FALSE, echo = F}
## knitr configuration: http://yihui.name/knitr/options#chunk_options
opts_chunk$set(comment = "", error= TRUE, warning = FALSE, message = FALSE,
               tidy = FALSE, cache = TRUE, echo = T,
               fig.width = 6, fig.height = 6)
```
Testing Test

========================================================

```{r}
set.seed(101)

# Read Data

size <- 50
my.data <- data.frame(x1 <- rpois(size,4),
                      x2 <- rnorm(size),
                      y <- x1*x2 + x1)

# Preprocess Data

names(my.data) <- c(paste0("x",1:(ncol(my.data)-1)),"y")

sapply(my.data, class)
my.data$x1 <- as.numeric(my.data$x1) # x1 becomes a numeric value
head(my.data,8)

# Execute ML methods

source("gapolyfuncs.R")
report <- make.report(my.data, population=20, iterations=20)

# See Results

coef(report$ga.model)
report$ga.error

boxplot(report$ga.error, 
        report$lm.error,
        report$svm.error,
        report$rpart.error,
        report$rf.error,
        report$citree.error,
        names=c("GA Poly", "Lin.Reg.", "SVM", "RPART", "Rnd.For.", "CI Tree"),
        ylab="RMSE")

# print(xtable(res.df), include.rownames = FALSE, floating = FALSE, type = "latex")
```