### 2015-08-11 texts/00-Previous/comparations/CoverinLetter.tex
Versão pré-final

\documentclass[11pt, oneside]{article}   	% use "amsart" instead of "article" for AMSLaTeX format
\usepackage{geometry}                		% See geometry.pdf to learn the layout options. There are lots.
\geometry{a4paper}                   		% ... or a4paper or a5paper or ... 
%\geometry{landscape}                		% Activate for for rotated page geometry
%\usepackage[parfill]{parskip}    		% Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}				% Use pdf, png, jpg, or eps§ with pdflatex; use eps in DVI mode
								% TeX will automatically convert eps --> pdf in pdflatex		
\usepackage{amssymb}

\title{Covering Letter}
\author{Francisco Coelho}
%\date{}							% Activate to display a given date or no date

\begin{document}
\maketitle
There is no conflict of interest with any of the suggested reviewers.

\hspace{7cm}\raggedright{\includegraphics{fcsig.pdf}}

\end{document}