### 2017-03-02 results/boxplot.stats/compute.boxplot.stats.Rmd
new review process
Changed project struture and updated files according to the 3rd review
process

```{r}
compute.box.plot  <- function(dataset) {
  df <- read.table(paste0("results/", dataset, "300_results_v2.txt"))

  res <- matrix(NA, ncol=5, nrow=ncol(df))

  for (i in 1:ncol(df))
    res[i,] <- boxplot.stats(df[,i])$stats

  res <- as.data.frame(res)
  names(res) <- c("min", "25%", "mean", "75%", "max")
  row.names(res) <- c("eprr.0.7", "eprr.0.8", "eprr.0.9", "epr", "lm", "svr", "rtree", "citree")
  res
}
```

```{r, fig.height=9}
datasets <- c("Abalone", "Auto-Mpg", "Housing", "Kinematics")

for(dataset in datasets) {
  print(dataset)
  df <- compute.box.plot(dataset)
  print(df)
  boxplot(t(df), main=dataset)
}
```