### 2017-03-02 results/svm_kernels/report.Rmd

```{r}
library(readr)
```

```{r, message=FALSE}
dataset <- "Abalone"

df <- read_csv2(paste0(dataset, ".csv"))

df$`polynomial deg=4` <- pmin(df$`polynomial deg=4`, 1.0)

boxplot(df$`polynomial deg=4`, df$linear, df$radial, names=c("poly", "linear", "radial"))
```

```{r, message=FALSE}
dataset <- "Auto-Mpg"

df <- read_csv2(paste0(dataset, ".csv"))

df$`polynomial deg=4` <- pmin(df$`polynomial deg=4`, 1.0)

boxplot(df$`polynomial deg=4`, df$linear, df$radial, names=c("poly", "linear", "radial"))
```

```{r, message=FALSE}
dataset <- "Housing"

df <- read_csv2(paste0(dataset, ".csv"))

df$`polynomial deg=4` <- pmin(df$`polynomial deg=4`, 1.0)

boxplot(df$`polynomial deg=4`, df$linear, df$radial, names=c("poly", "linear", "radial"))
```

```{r, message=FALSE}
dataset <- "Kinematics"

df <- read_csv2(paste0(dataset, ".csv"))

df$`polynomial deg=4` <- pmin(df$`polynomial deg=4`, 1.0)

boxplot(df$`polynomial deg=4`, df$linear, df$radial, names=c("poly", "linear", "radial"))

```